---
title: "Model Specification"
author: "Carlos Calvo-Hernandez"
date: "5/7/2020"
output: 
  pdf_document:
    toc: TRUE
    number_sections: TRUE
    fig_width: 5.5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(deSolve)
source("data_input.R")
```

# Basic R Model of the U.S. Budget

## Basic Model of Economic Growth

$$
\text{GDP}(t) = \text{GDP}(t-1)*\partial \text{GDP}(t-1)
$$
where,

$$
 \partial \text{GDP}(t-1) = (\partial \text{ MFP} + \partial L * \text{ GDP}_{shareL} + \partial K * \text{ GDP}_{shareK}) + \partial \text{ GDP}(\text{debt}) + \sum_i \partial \text{ GDP}(\text{shock}_i, t)
$$



multifactor productivity
$$
\partial \text{ MFP} = \partial \text{ MFP}(trend)*\partial \text{ MFP}(K) + \partial \text{ MFP}(\text{PublicInvestment})
$$


Labor
$$
\partial L = L(\text{trend}) * [1+\partial L(Tax)+\partial L(Infrastructure)]
$$

Capital
$$
\partial K = K(\text{trend})*[1 +\partial K(Tax)+\partial K(Infrastructure)]
$$

with, $\text{ GDP}_{shareL} +  \text{ GDP}_{shareK} =1$




### Basic model for GDP Growth

For 

$$
Y_t = Y_{t-1}* (1+y_{t-1})
$$
where
$$
y_t = [\alpha_t + l_t * (Y_L/Y) + k_t * (Y_K/Y)] + \nu_{debt(t)} + \nu_{shock(t)}
$$

## Growth Model

### Labor parameters

```{r}
pop0 <- 140L
dl <- rep(NA, 76)
dl[1] <- 0.084
for (t in 2:76) {
  if (t != 2) {
    dl[t] <- (population[t - 1] / population[t - 2]) - 1
  } else {
    dl[t] <- (population[t] / pop0) - 1
  }
}
```

### Capital parameters

```{r}
dk <- rep(NA, 76)
dk[1] <- 0.0181
for (t in 2:76){
  dk[t] <- dk[1]*(1 + kTax[t] + kInf[t])
}
```

### Parameter Initialization and Initial Conditions

```{r}

k <- rep(NA, 76)
y <- rep(NA, 76)
g <- rep(NA, 76)
a <- rep(NA, 76)
lPol <- rep(NA, 76)
kPol <- rep(NA, 76)
aPol <- rep(NA, 76)

Ly <- 0.6
Ky <- 0.4
a1 <- 0.012

k[1] <- 0.03 
y[1] <- 15000
g[1] <- 0.018 #y_t in the model
a[1] <- 0.0112

li <- 0.2
ki <- 0.1
```

### Model Specification

```{r model_growth}
for (t in 2:76){
  #aPol[t] <- lPol[t] + kPol[t]
  a[t] <- a[t-1]*(1+dk[t]) #+ aPol[t-1]
  g[t] <- a[t] + dl[t]*Ly + dk[t]*Ky 
  y[t] <- y[t-1]*(1+g[t-1])
}
```


```{r}
plot.ts(y)
```

```{r}
plot.ts(g)
```

```{r}
coeff <- sqrt(sum(y^2))
df <- tibble(growth = y, mfp = g, time = (1:76))
ggplot(data = df, aes(x = time))+
  geom_line(aes(y = growth))+
  geom_line(aes(y = mfp*coeff)) +
  scale_y_continuous(name = "GDP",
                     sec.axis = sec_axis(~./coeff, name =  "MFP"))
```

```{r}
plot.ts(a)
```


## Setting up the debt model

### Fiscal Policy Influences on Growth Rate

$$
\begin{aligned}
l_{\text{policy}(t)} &= l_{\text{tax personal}(t)} + l_{\text{infrastructure}(t)}\\
k_{\text{policy}(t)} &= k_{\text{tax coporate}(t)} + k_{\text{infrastructure}(t)}\\
a_{\text{policy}(t)} = a_{\text{defense}(t)} + a_{\text{education}(t)} &+ a_{\text{health}(t)} + a_{\text{infrastructure}(t)} + a_{\text{research}(t)} + a_{\text{social}(t)}
\end{aligned}
$$

$$
a_{\text{target policy}(t)} = \frac{da}{d\tau_{\text{target policy}(t)}}* (\tau_{\text{target policy}(t)})
$$
where $\frac{da}{d\tau_{\text{target policy}(t)}}$ is the elasticity of the rate of growth of multifactor productivity (or of labor or capital as specified) to the expenditure for the target policy, and $\tau_{\text{target policy}(t)}$ is the expenditure for the target policy

$$
\nu_{\text{debt}(t)} = \left[ \frac{d\nu}{d (\text{Debt})/Y}  \right]*\left(\frac{\text{Debt}_t}{Y_t}\right)
$$


where $\frac{d\nu}{d (\text{Debt})/Y}$ is the elasticity of the economic growth rate to the debt-to-GDP ratio, and $\frac{\text{Debt}_t}{Y_t}$ is the debt-to-GDP ratio in time $t$.

$$
\text{Debt}_t = \text{Debt}_{t-1} + \text{Balance}_{t-1}
$$

where
$$
\text{Balance}_t = \text{Taxes}_{t} + \text{Expenditures}_{t}
$$
and,

$$
\text{Expenditures}_t = \text{Mandatory}_t + \text{Discretionary}_t + \text{Interest}_t
$$

where,
$$
\text{Mandatory}_t = \text{SocialSecurity}_t + \text{Health}_t + \text{Other}_t
$$

$$
\text{Discretionary}_t = \text{Defense}_t + (\text{Education}_t + \text{Infrastructure}_t + \text{Research\&Development}_t)
$$



### Debt Model Specification

#### Interest Rate Parameters

```{r interest_det}
# Interest Rate set up
intRate <- rep(NA, 76)
intRateSet <- values$Experimental[28] #"FIAT" or "MARKET"
intRate[1] <- 0.027 # B166
intRateTrend <- 0.058 # B167
intThreshold <- 1.5 # B168
dIntdy <- as.numeric(values$Experimental[30])/100 # dInterest/dGDP  B169
dIntdDebtY <- as.numeric(values$Experimental[31])/100 # dInterest/d(Debt/GDP) B170
intConstant <- values$Experimental[32]/100 # B171

```

#### Parameter Initialization

```{r series_bud}
debt <- rep(NA, 76)
bal <- rep(NA, 76)
tax <- rep(NA, 76)
exp <- rep(NA, 76)
interest <- rep(NA, 76)
discrete <- rep(NA, 76)
mand <- rep(NA, 76)
personal <- rep(NA, 76)
corp <- rep(NA, 76)
otherTax <- rep(NA, 76)
otherExp <- rep(NA, 76)
income <- rep(NA, 76)
fica <- rep(NA, 76)
defense <- rep(NA, 76)
RD <- rep(NA, 76)
educ <- rep(NA, 76)
infr <- rep(NA, 76)
health <- rep(NA, 76)
social <- rep(NA, 76)
```


#### Tax Specification

##### Baseline data

```{r echo = FALSE}
knitr::kable(budgetLines)
```
##### Tax Revenue

```{r}
# Tax Revenue
taxes <- function(type, k) { # personal, corporate and Other taxes
  for (t in 1:76) {
    type[t] <- y[t] * (budgetLines$trend[k] * (1 + budgetLines$pol[k])) + (1 + budgetLines$adap[k])
  }
  return(type)
}

income <- taxes(income, 1)
fica <- taxes(fica, 2)
corp <- taxes(corp, 3)
otherTax <- taxes(otherTax, 4)
otherExp <- taxes(otherExp, 7)
defense <- taxes(defense, 8)
RD <- taxes(RD, 9)
educ <- taxes(educ, 10)
infr <- taxes(infr, 11)
# Health and social expenditures as %GDP (need to change as cost of beneficiaries)
health <- taxes(health, 6)
social <- taxes(social, 5)
```



```{r }
 #Health and social protection policies
 #health[t] <- (healthTrend*(1 + healthPol)*beneficiary)*(1 + healthAdap[t])
 #social[t] <- (socialTrend *(1 + socialPol)*beneficiary)*(1 + socialAdap[t])
```


#### Endogenous Interest

```{r}
# Endogenous interest
if (intRateSet == 1) { # 1 = FIAT and 0 = MARKET
  for (i in 1:length(intRate)) {
    intRate[i] <- intRateTrend
  }
} else {
  for (j in 1:length(intRate)) {
    intRate[j] <- 0.015 + 0.45 * g[j - 1] * (0.02(debt[j - 1] / y[j - 1]))
  }
}
  
```

#### Initial Conditions for Revenues and Expenditures

```{r}
# Initial conditions

debt[1] <- 10350
interest[1] <- debt[1]*intRate[1]
discrete[1] <- defense[1] + (educ[1] + infr[1] + RD[1]) #discretionary expenditure
mand[1] <- social[1] + health[1] + otherExp[1] # Mandatory Expenditures
exp[1] <- mand[1] + discrete[1] + interest[1] # Total Expenditures

personal[1] <- income[1] + fica[1]
tax[1] <- personal[1] + corp[1] + otherTax[1] # Taxes

bal[1] <- tax[1] - exp[1] # GDP Balance
```


#### Model Specification

```{r rev-tax-model}

for (t in 2:76){
interest[t] <- debt[t-1]*intRate[t] # Interest Expenditures

discrete[t] <- defense[t] + (educ[t] + infr[t] + RD[t]) # Discretionary Expenditure
mand[t] <- social[t] + health[t] + otherExp[t] # Mandatory Expenditures
exp[t] <- mand[t] + discrete[t] + interest[t] # Total Expenditures

personal[t] <- income[t] + fica[t] # Personal taxes
tax[t] <- personal[t] + corp[t] + otherTax[t] # Taxes

bal[t] <- tax[t] - exp[t] # GDP Balance
debt[t] <- debt[t-1] + bal[t-1]
}
```

#### Expenditure Plot

```{r fig.align='center'}
plot.ts(exp)
```


```{r}
debtGDPratio <- debt/y
```


#### Debt to GDP Ratio Plot

```{r fig.align='center'}
plot.ts(debtGDPratio)
```
```{r}
plot.ts(debt)
```


### Debt Ratio and Adaptive Policies Specification
```{r}
dGDPdebt <- rep(NA, 76)
```



```{r}
# Changes in Labor and Capiital across the budget
## include dGDP(debt) into dGDP (g[t]) calculation

dGDPdebt[1] <- min(0, -({debtGDPratio[1] - economyDashboard$Value[4]}*economyDashboard$Value[3]))

```

```{r}
for (t in 2:76) {
  dGDPdebt[t] <- min(0, -({debtGDPratio[t - 1] - economyDashboard$Value[4]}*economyDashboard$Value[3]))
}
```

```{r}
plot.ts(dGDPdebt)
```

```{r}
aDebt <- a
gDebt <- g
yDebt <- y
```


```{r gdp-debt}
### GDP with debt
for (t in 2:76){
  #aPol[t] <- lPol[t] + kPol[t]
  aDebt[t] <- a1*dk[t] #+ aPol[t-1]
  gDebt[t] <- {a[t] + dl[t]*Ly + dk[t]*Ky} + dGDPdebt[t]
  yDebt[t] <- y[t-1]*(1+g[t-1])
}

```


```{r}
plot.ts(yDebt)
```

```{r}
debtGDPratio2 <- debt/yDebt
plot.ts(debtGDPratio2)
```


## Changing Population module

```{r}
# Parameter initialization
GDP <- y
dPopdGDP <- rep(NA, 76)
dPop <- rep(NA, 76)

dGDP <- rep(NA, 76)
dGDP[1] <- 0

# GDP <- seq(from = 15000, to = 9500, length.out = 76) ## to check for negative dGDP behavior


# Functions

#dl[t] <- dPop[t] #dl as in the GDP calculation or input for the model?

dPop <- dl

PopGDPRatio <- 0.5

#population[t] <- population[t - 1]*(1 + dPop[t])

#dPop[t] <- dPop[t - 1] + dPopdGDP[t]

for (t in 2:76)  {
  
  dGDP[t] <- GDP[t]/GDP[t - 1] - 1
  
    if (dGDP[t - 1] >= 0) {
    dPopdGDP[t] <- 0
  } else {
    dPopdGDP[t] <- (PopGDPRatio)*dGDP[t - 1] # this is a circular equation, need to figure out if dPop and dGDP are on timestep t or t-1
  }
  
  dPop[t] <- dPop[t - 1] + dPopdGDP[t]
  
  population[t] <- population[t - 1]*(1 + dPop[t])
  
    }


```


```{r}
plot.ts(population)
```


```{r}
plot.ts(dPop)
```
```{r}
plot.ts(dPopdGDP)
```


```{r}
plot.ts(dGDP)
```



















